### 2022-05-31 README.Rmd
add actual readme

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Why does preregistration increase the persuasiveness of evidence? A Bayesian rationalization.

```{r, echo=FALSE, results='asis'}
cat(readLines('abstract.md'))
```