### 2022-09-12 README.Rmd

---
output: github_document
bibliography: references.bib
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Why does preregistration increase the persuasiveness of evidence? A Bayesian rationalization.

```{r abstract, echo=FALSE, results='asis'}
cat(readLines('abstract.md'))
```

## [🗎 PDF of draft 🗎](https://aaronpeikert.github.io/bayes-prereg/manuscript.pdf)

## [Slides for Talk](https://aaronpeikert.github.io/bayes-prereg/presentation.html)

## References